Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid

### 2016-02-17 Article.tex
modifs intro

The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures to name a few, sketch a new power system. The rigid top-down centralized architecture of today’s power grids is expected to evolve in more distributed and flexible systems. Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition. Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages.

In this paper, we consider agents in the distribution network that own both renewable distributed energy resources (DER) and loads. We will refer to these agents as prosumers \cite{Rathnayaka2012} in the future. If a prosumer produces more than he consumes, he might be willing to sell power. We thus assume the existence of an energy market where producers post a production capacity for an upcoming period, and buyers decide what bids to purchase. When the production is stochastic, as it can be with DER, this approach gets more complicated and risk should be quantified and taken into account. Rather than having a risk dependant price, we propose here a safer and simpler model. We assume the existence of a market operator who constrains the market entrance. A key point of this work is to merge the interests of the market operator with the prosumers in a single utility function. While the prosumers intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply.

If a buyer purchases a contract, the corresponding producer has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates. Since prosumers use renewables, we are concerned with the probability that the production falls below the contract value. In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low. Single prosumers might be authorized to enter such market, but it is unlikely that a single agent can propose enough power with enough reliability.

Therefore, we use a coalitional framework in order to model aggregators of prosumers. It is known that diversification of the assets is a way of minimizing risks when building a portfolio. We thus expect more stable and predictable energy productions for aggreagtors than for single agents. Nevertheless, all coalitions are not equally stable or productive. Attention should thus be paid to the aggregation step. Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents. This is a central topic of this paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?

>>>>

%The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures to name a few, sketch a new power system.

%Bidirectional power flows, storage, and distributed production are characteristics of smart grid systems. Nodes that were simply pure loads yesterday could behave tomorrow as generators or loads depending on weather conditions.

%Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition. Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages.

In this paper, we consider agents in the distribution network that own both renewable distributed energy resources (DER) and loads. We will refer to these agents as prosumers \cite{Rathnayaka2012} in the future. If a prosumer produces more than he consumes, he might be willing to sell power. We thus assume the existence of an energy market where producers post a production capacity for an upcoming period, and buyers decide what bids to purchase. When the production is stochastic, as it can be with DER, this approach gets more complicated and risk should be quantified and taken into account. Rather than having a risk dependant price, we propose here a safer and simpler model. We assume the existence of a market operator who constrains the market entrance.

As single agents might not be productive or reliable enough, aggregators managing large portfolios of prosumers are more likely to enter such markets. A key point of this work is to merge the interests of the market operator and aggregators in a single utility function. While the aggregators intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply.

If a buyer purchases a contract, the corresponding aggregator has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates. Since prosumers use renewables, we are concerned with the probability that the production falls below the contract value. In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.

In this paper, we decided to use a coalitional framework in order to model aggregators. It is known that diversification of the assets is a way of minimizing risks when building a portfolio. We thus expect more stable and predictable energy productions for aggreagtors than for single agents. Nevertheless, all coalitions are not equally stable or productive. Attention should thus be paid to the aggregation step. Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents. This is a central topic of this paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
